---
title: "Project Draft"
author: "Nischal Bhandari"
date: "06/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages

```{r load-packages, message=FALSE}
library(tidyverse) 
library(skimr)
library(datasauRus)
library(dplyr)
```

### Load data

```{r load-data, message = FALSE}
heart <- read_csv("data/heart.csv")
```


Ex. 1. 

### Understanding our dataset

There are `r ncol(heart)` variables and `r nrow(heart)` observations in our dataset. 

We will print out the summary of our dataset. 

```{r echo=FALSE}
heart %>%
  skim_without_charts()
```


Ex. 2

We will plot to visualize the relationship between `age` and `resting blood pressure` of a patient with a heart disease.

```{r echo = FALSE}
ggplot(heart, mapping = aes(x = age, y = trestbps))+
  geom_point()+
  geom_line(color = "violetred") +
  labs(
  x = "Age/years", 
  y = "Resting Blood Pressure",
  title = "Resting Blood Pressure among patients of different ages"
)


```


It can not be inferred clearly the relationship between `age` and `resting blood pressure`, but generally aged patients tend to have more `resting blood pressure`. 


### Summarising the `resting blood pressure` 

```{r echo = FALSE}
heart %>%
  select(trestbps) %>%
  summary()
```


The minimum resting blood pressure is 94.0, naximum is 200.0, and mean is 131.6.


Ex. 3.

We will now calculate summary for male and female patients.

First of all, we will mutate `sex` variable. 

```{r}
heart <- heart %>%
mutate(sex = case_when(
  sex== 0 ~ "Female",
  sex == 1 ~ "Male") 
)
  
heart
```

Now, we have categorized gender of patients as a character variable. 

We will create boxplot for `age`, `resting blood pressure`, `gender` variables. 

```{r echo = FALSE}
ggplot(heart, aes(x=age, y=trestbps, fill = sex)) + geom_boxplot() + 
   
  labs(
    x = "Age/years", 
    y = "Resting blood pressure"
  )
  
```

Ex. 4

We will filter patients that are above 50 to visualize the relationship between their `age` and `blood cholestoral`. 

```{r echo = FALSE}
heart_old <- heart %>%
  filter(age > 50)

ggplot(heart_old,mapping = aes(x = age, y= chol))+
  facet_wrap(.~sex) +
  geom_point()+
  geom_smooth(color ="red")+
  labs(
x = "Age/year",
y = "serum cholestoral in mg/dl",
title = "Old aged patients and Cholestoral",
subtitle = "by gender")
```


Generally, the 'blood cholesterol` in female is low as their age increases, but in case of male, their cholesterol level generally rises with the increase in their age. 

Ex. 5

We'll count the cases of excercise-induced angina among heart patients. 

### Mutating the variable first. 

```{r echo=FALSE}
heart <- heart %>%
  mutate(exang = case_when(
    exang == 1 ~ "Yes",
    exang  == 0 ~ "NO")
  )
heart
```


Now, we will count the number of exercise-induced angina. 

```{r}
heart %>%
  count(exang, sort = TRUE)
```

 So 204 patients didn't have exercise-induced angina while 99 had. 


Ex. 6

We will love to see if such induced angina was more prevalent among male. 

```{r echo FALSE}
ggplot(heart, mapping = aes(x = age, y = exang))+
geom_jitter(color="red")+
facet_wrap(.~sex)+
labs(
title = "Prevalence of exercise-induced angina among heart patients",
subtitle = "by gender"
)
```

Comparatively, less number of women with heart diseases have exercise-induced angina than man.


Ex. 7

Now, we will counnt the distribution of `chest pain` among our patients. 

```{r echo = FALSE}
heart %>%
  count(cp, sort = TRUE)
```

Majority i.e. 143 patients had `type-0` chest pain while 87 had `type-2`, 50 `type-1` and 23 `type-3` which represent least number of patients.  
