---
title: "Project Draft"
author: "Nischal Bhandari"
date: "06/12/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages

```{r load-packages, message=FALSE}
library(tidyverse) 
library(skimr)
library(datasauRus)
library(dplyr)
library(tidymodels)
```

### Load data

```{r load-data, message = FALSE}
heart <- read_csv("data/heart.csv")
```

### Introduction

 The dataset I used includes observations among potential heart disease patients. The observations are done in the following aspects of patients: 
 
1.  age

2. sex

3. chest pain type (4 values)

4. resting blood pressure

5. serum cholesterol in mg/dl

6. fasting blood sugar > 120 mg/dl

7. resting electrocardiographic results (values 0,1,2)

8. maximum heart rate achieved

9. exercise induced angina

10. oldpeak = ST depression induced by exercise relative to 
rest

11. the slope of the peak exercise ST segment

12. number of major vessels (0-3) colored by flourosopy

13. thal: 3 = normal; 6 = fixed defect; 7 = reversible defect

14. target (if patient have heart disease)


### Heart dataset 
This dataset is claimed to be frequently used by ML researchers. It is the work of Cleaveland datase that is contributed by: 

1. Hungarian Institute of Cardiology. Budapest: Andras Janosi, M.D.

2. University Hospital, Zurich, Switzerland: William Steinbrunn, M.D.

3. University Hospital, Basel, Switzerland: Matthias Pfisterer, M.D.

4. V.A. Medical Center, Long Beach and Cleveland Clinic Foundation: Robert Detrano, M.D., Ph.D.

I obtained this dataset from this link: https://www.kaggle.com/ronitf/heart-disease-uci 


### Objective:

My objective on this project is to tinker around the potential links between various bodily, metabolic conditions rather than reach a specific conclusions, so I potentially won't have much hypotheses. 

Ex. 1. 



### Understanding our dataset

There are `r ncol(heart)` variables and `r nrow(heart)` observations in our dataset. 

We will print out the summary of our dataset. 

```{r echo=FALSE}
heart %>%
  skim_without_charts()
```


Ex. 2

We will plot to visualize the relationship between `age` and `resting blood pressure` of a patient.

```{r echo = FALSE}
ggplot(heart, mapping = aes(x = age, y = trestbps))+
  geom_point()+
  geom_line(color = "violetred") +
  labs(
  x = "Age/years", 
  y = "Resting Blood Pressure",
  title = "Resting Blood Pressure among patients of different ages"
)


```


It can not be inferred clearly the relationship between `age` and `resting blood pressure`, but generally aged patients tend to have more `resting blood pressure`. 


### Summarising the `resting blood pressure` 

```{r echo = FALSE}
heart %>%
  select(trestbps) %>%
  summary()
```


The minimum resting blood pressure is 94.0, naximum is 200.0, and mean is 131.6.


Ex. 3.

We will now calculate summary for male and female patients.

First of all, we will mutate `sex` variable. 

```{r echo = FALSE}
heart <- heart %>%
mutate(sex = case_when(
  sex== 0 ~ "Female",
  sex == 1 ~ "Male") 
)
  
heart
```

Now, we have categorized gender of patients as a character variable. 

We will create boxplot for `age`, `resting blood pressure`, `gender` variables. 

```{r echo = FALSE}
ggplot(heart, aes(x=age, y=trestbps, fill = sex)) + geom_boxplot() + 
   
  labs(
    x = "Age/years", 
    y = "Resting blood pressure"
  )
  
```

Ex. 4

We will filter patients that are above 50 to visualize the relationship between their `age` and `blood cholestoral`. 

```{r echo = FALSE}
heart_old <- heart %>%
  filter(age > 50)

ggplot(heart_old,mapping = aes(x = age, y= chol))+
  facet_wrap(.~sex) +
  geom_point()+
  geom_smooth(color ="red")+
  labs(
x = "Age/year",
y = "serum cholestoral in mg/dl",
title = "Old aged patients and Cholestoral",
subtitle = "by gender")
```


Generally, the 'blood cholesterol` in female is low as their age increases, but in case of male, their cholesterol level generally rises with the increase in their age. 

Ex. 5

We'll count the cases of excercise-induced angina among heart patients. 

### Mutating the variable first. 

```{r echo = FALSE}
heart <- heart %>%
  mutate(exang = case_when(
    exang == 1 ~ "Yes",
    exang  == 0 ~ "NO")
  )
heart
```


Now, we will count the number of exercise-induced angina. 

```{r}
heart %>%
  count(exang, sort = TRUE)
```

 So 204 patients didn't have exercise-induced angina while 99 had. 


Ex. 6

We will love to see if such induced angina was more prevalent among male. 

```{r echo FALSE}
ggplot(heart, mapping = aes(x = age, y = exang))+
geom_jitter(color="red")+
facet_wrap(.~sex)+
labs(
title = "Prevalence of exercise-induced angina among heart patients",
subtitle = "by gender"
)
```

Comparatively, less number of women with heart diseases have exercise-induced angina than man.


Ex. 7

Now, we will counnt the distribution of `chest pain` among our patients. 

```{r echo = FALSE}
heart %>%
  count(cp, sort = TRUE)
```

Majority i.e. 143 patients had `type-0` chest pain while 87 had `type-2`, 50 `type-1` and 23 `type-3` which represent least number of patients.  


Ex. 8.

We will try to model `thalach` which is 
maximum heart rate achieved as dependent on `age`, `the resting blood pressure`, and `Cholesterol level`. 

```{r echo = FALSE}
m_heart <- lm (thalach ~ age + chol + trestbps, data = heart)
tidy(m_heart)
```

The model can be expressed in a formula as: 

$predicted_maximum_heart_rate$ = $190 +  0.0329*chol - 1.09*age + 0.0849*trestbps$


Ex. 9

We will check if our model represent an accurate interaction between variables. 

```{r echo=FALSE}
glance(m_heart)$r.squared
glance(m_heart)$adj.r.squared
```

The R squared value is 0.1686218 while adjusted R squared value is 0.1602802. These values are pretty low, so this model doesn't represent good interaction between the response and independent variables. 
